---
title: "labbook_151222"
author: "Steve Harris"
date: "30 Dec 2015"
output: "html_document"
---
Approach to the analysi fo the European septic shock data

```{r include=FALSE}

rm(list=ls(all=TRUE))

# Set up and load data

require(Hmisc)
require(assertthat)
require(gmodels)
require(descr)
require(data.table)
require(XLConnect)
require(ggplot2)
require(arm) 			# for display

# Pander options
require(pander)
panderOptions('table.split.table', Inf)
panderOptions('table.alignment.default', 'right')
panderOptions('table.style', 'rmarkdown')

load(file='../data/cleaned.Rdata')
load(file='../data/clean_long.Rdata')

wdt.original <- wdt
wdt$sample_N <- 1
wdt$all <- 1
wdt[, all := factor(all, label=("All patients"))]
str(wdt)
assert_that(nrow(wdt)==736)
names(wdt)

# Mosaic plot with ggplot function
ggMMplot <- function(var1, var2){
  require(ggplot2)
  levVar1 <- length(levels(var1))
  levVar2 <- length(levels(var2))

  jointTable <- prop.table(table(var1, var2))
  plotData <- as.data.frame(jointTable)
  plotData$marginVar1 <- prop.table(table(var1))
  plotData$var2Height <- plotData$Freq / plotData$marginVar1
  plotData$var1Center <- c(0, cumsum(plotData$marginVar1)[1:levVar1 -1]) +
    plotData$marginVar1 / 2

  ggplot(plotData, aes(var1Center, var2Height)) +
    geom_bar(stat = "identity", aes(width = marginVar1, fill = var2), col = "White") +
    scale_fill_brewer(type="seq", palette="YlOrRd") +
    geom_text(aes(label = as.character(var1), x = var1Center, y = 1.05, size=0.5)) +
    theme_minimal()
}


```

We have `r nrow(wdt)` cases from `r length(unique(wdt$hosp))` sites.

```{r, echo=FALSE, result=TRUE}
wdt[,.(n=.N,pct=round(.N/nrow(wdt)*100)), by=hosp]
```

We set relatively strict and clear inclusion criteria so the cases should be homogeneous with similar characteristics _across_ sites. In particular, mortality, physiology (heart rate and blood pressure), and use of noradrenaline should not differ much.

But this is not true for mortality at either unit or hospital level.

### Mortality

```{r, 'plot_mort', echo=FALSE, result=TRUE}
mosaicplot(hosp ~ factor(mort.itu, levels=c(0,1), labels=c("Alive", "Dead")),
	color=c("grey80", "grey0"), data=wdt, border=FALSE)
title(main="", ylab="ITU mortality (%)")


t <- with(wdt, table(mort.itu.order, factor(mort.itu, levels=c(0,1), labels=c("Alive", "Dead"))))
mosaicplot(t, color=c("grey80", "grey0"), border=FALSE)
title(main="", xlab="Hospital", ylab="ITU mortality (%)")
```

So if mortality differs, do we see a corresponding pattern in the physiology? Let's examine heart rate for a starter.

### Heart rate

```{r, echo=FALSE, result=TRUE}
print("Heart rate at hour 1")
summary(wdt$hr.1)
print("Heart rate at hour 24")
summary(wdt$hr.24)

require(ggplot2)
str(wdt.long)
p <- ggplot(data=wdt.long, aes(y=hr, x=mort.itu.order))
p + geom_boxplot(notch=TRUE) +
	facet_grid(.~time) +
	coord_cartesian(ylim=c(40,160))
```

So Leuven has high mean HR and Anconca a low which gives the impression that there is an inverse relationship with mortality. But if you remove these two sites then it is less clear.


```{r, echo=FALSE, result=TRUE}
table(wdt$hosp)
p <- ggplot(data=wdt.long[!(hosp %in% c("Ancona", "Leuven"))],
		aes(y=hr, x=mort.itu.order))
p + geom_boxplot(notch=TRUE) +
	facet_grid(.~time) +
	coord_cartesian(ylim=c(40,160))
```

### Mean arterial pressure

Let's check and look at Mean Arterial Pressure.

```{r, echo=FALSE, result=TRUE}
print("MAP at hour 1")
summary(wdt$map.1)
print("MAP at hour 24")
summary(wdt$map.24)

p <- ggplot(data=wdt.long, aes(y=map, x=mort.itu.order))
p + geom_boxplot(notch=TRUE) +
	facet_grid(.~time) +
	coord_cartesian(ylim=c(40,120))
```

So this looks more interesting? Higher MAPs are seen in the sites with the highest ITU mortality at both hour 1 and 24.

Is this driven by noradrenaline use?

### Noradrenaline use

```{r, echo=FALSE, result=TRUE}
print("Noradrenaline at hour 1")
summary(wdt$ne.1)
print("Noradrenaline at hour 24")
summary(wdt$ne.24)

p <- ggplot(data=wdt.long, aes(y=ne, x=mort.itu.order))
p + geom_boxplot(notch=TRUE) +
	facet_grid(.~time) +
	coord_cartesian(ylim=c(0,1))
```

It looks like MAP and noradrenaline dose are correlated, but Paris seems to be an exception here as the noradrenaline does is high but the mortality is lowish.

Is there any chance that this might be due to different methods for calculating noradrenaline dose (dividing by ideal vs actual body weight?)

Now have a look at fluid balance and sedation practice.

### Fluid balance

```{r, echo=FALSE, result=TRUE}
print("Fluid balance in 1st 24 h")
summary(wdt$fb.24)

p <- ggplot(data=wdt.long, aes(y=fb, x=mort.itu.order))
p + geom_boxplot(notch=TRUE) +
	facet_grid(.~time) +
	coord_cartesian(ylim=c(-2000,8000))
```

### Sedation scores

```{r, echo=FALSE, result=TRUE}
print("Sedation scores by site")
with(wdt, table(as.factor(sedation.1), mort.itu.order))

describe(wdt$sedation.1)
describe(wdt$sedation.24)

wdt[, sed1 := factor(round(sedation.1 / 2),
	labels=c("0", "+", "++", "+++", "++++"))]
wdt[, sed24 := factor(round(sedation.24 / 2),
	labels=c("0", "+", "++", "+++", "++++", "+++++"))]


with(wdt[!is.na(sed1)], ggMMplot(mort.itu.order, sed1))
with(wdt[!is.na(sed24)], ggMMplot(mort.itu.order, sed24))

```

I note that Paris uses a lot of sedation which perhaps explains the higher noradrenaline requirements?
