---
title: "labbook_151222"
author: "Steve Harris"
date: "22 Dec 2015"
output: "html_document"
---
Notes on analysis of the European septic shock data

```{r include=FALSE}

rm(list=ls(all=TRUE))

# Set up and load data

require(Hmisc)
require(assertthat)
require(gmodels)
require(descr)
require(data.table)
require(XLConnect)
require(ggplot2)
require(arm) 			# for display

# Pander options
require(pander)
panderOptions('table.split.table', Inf)
panderOptions('table.alignment.default', 'right')
panderOptions('table.style', 'rmarkdown')

load(file='../data/cleaned.Rdata')

wdt.original <- wdt
wdt$sample_N <- 1
wdt$all <- 1
wdt[, all := factor(all, label=("All patients"))]
str(wdt)
assert_that(nrow(wdt)==736)
names(wdt)
attach(wdt)

```

## Tasks

- [X] import and clean data
- [X] prepare patient level version of Table 1

## Notes


### Quick inspect of patient level data

We have `r nrow(wdt)` cases available to analyse.

```{r, inspect_hosp, echo=FALSE, result=TRUE}
pander(table(hosp))
```

```{r, inspect_table1, echo=FALSE, result=TRUE}
load("../outputs/tables/table1_all.RData")
t1.wide.summ[,c(1:2,4:5),with=FALSE]
```

Primary outcome is mortality, and we have both ITU and hospital mortality with minimal missing data.


```{r, 'descr mortality', echo=FALSE, result=TRUE}
print("ITU mortality")
t <- wdt[, .(.N,"%"=round(100*.N/nrow(wdt))), by=.("ITU mortality"=mort.itu)]
pander(t)

print("Hospital mortality")
t <- wdt[, .(.N,"%"=round(100*.N/nrow(wdt))), by=.("Hospital mortality"=mort.hosp)]
pander(t)
```

So how much of these differences can be ascribed to baseline characteristics? In other words, there are two ways of looking at this: either the patients are different, and if we had similar patients then we would see similar outcomes; or, the treatment is different and this is the cause for differences in outcome.

```{r, 'plot_mort', echo=TRUE}
mosaicplot(hosp ~ mort.itu,
	color=c("grey80", "grey0"), border=FALSE)

# Now let's facet by site
wdt$hosp.ne.24 <- NULL
wdt[, hosp.ne.24 := quantile(ne.24, 0.5, na.rm=TRUE), by=hosp]
wdt[,.(min(hosp.ne.24)), by=hosp]
wdt[, hosp.ne.24 := factor(hosp.ne.24, labels=c(LETTERS[1:8]))]
str(wdt$hosp.ne.24)
with(wdt, mosaicplot(hosp.ne.24 ~ mort.itu,
	color=c("grey80", "grey0"), border=FALSE))

describe(ne.1)
# Scale 10* for interpretability
wdt[, ne.1s := ne.1*10]
wdt[, ne.24s := ne.24*10]

m <- glm(mort.itu ~ ne.24s, data=wdt, family=binomial(link="logit"))
display(m)
wdt[, yhat := invlogit(predict(m))]
mean(as.numeric(wdt$mort.itu)-1, na.rm=TRUE)
ggplot(data=wdt, aes(y=yhat,x=ne.24s)) +
	geom_point() +
	geom_smooth() +
	coord_cartesian(y=c(0,1), x=c(0,10))

```



```{r include=FALSE}
detach(wdt)
```










