---
title: "scratch_results"
author: "Steve Harris"
date:	"7 Mar 2016"
output: md_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls(all=TRUE))
source(file="../prep/load.R")
load(file='../data/cleaned.Rdata')
source(file="../prep/strobe.R")
source(file="../prep/prep_vars.R")
source(file="../share/functions4rmd.R")
```

# Methods

## Patients

We recruited one hundred consecutive admissions with septic shock from eight critical care units across Europe. Patients were eligible if they were locally diagnosed and treated for infection, required a vasopressor infusion after 24 hours of stabilisation in the critical care unit, and had not received beta-blocker therapy during the same time period.

<!-- Morelli criteria to compare in table

- age < 18
- beta-blockers prior to randomisation
- cardiac index < 2.2 with high LAP
- significant valvular heart disease
 -->

## Definitions

- need to describe vasopressor and sedation scores

## Statistical analaysis	



# Results
```{r strobe, include=FALSE, echo=FALSE}

los.itu.0    <- ff.np(exclude.los.itu.0)
rx.betablock <- ff.np(exclude.rx.betablock)
low.ne.24    <- ff.np(exclude.ne.24)
hosp.n       <- ff.mediqr(N, data=tdt[,.N,by=hosp])
```

We screened `r nrow(wdt[include==1])` consecutive admissions with septic shock from `r length(unique(wdt$hosp))` hospitals in 4 European countries. We excluded `r los.itu.0$n[2]` patients (`r los.itu.0$p[2]`) who were discharged within 24 hours (one survivor, one non-surivor); `r rx.betablock$n[2]` patients (`r rx.betablock$p[2]`) who received beta-blocker therapy during the first 24 hours; and, `r low.ne.24$n[2]` patients (`r low.ne.24$p[2]`) where noradrenaline infusions had been weaned by 24 hours. The final cohort included `r nrow(tdt)` patients with each hospital contributing a median of `r hosp.n$q50` patients (IQR `r hosp.n$iqr`).

## Patient characteristics

```{r patients, include=FALSE, echo=FALSE}
# source(file="../analysis/table1_all.R")
```
<!-- see Table 1 -->

Vasopressor support for septic shock commenced on the first ICU day for 311 (75%) patients, and pneumonia was the commonest cause of infection (223 patients, 46%). The median SOFA score at 24 hours was 10 (IQR 8 to 12). The majority of patients (424, 87%) were mechanically ventilated, and 140 (29%) had already started renal replacement therapy (RRT) at 24 hours. The overall ICU and hospital mortalities were 41% (198 deaths), and 48% (233 deaths) respectively.

## Variation in practice between sites

- blood pressure targets
- noradrenaline dosing
- sedation
- fluid management
- mortality

## Variation in mortality between sites

a
## Model 24 hour noradrenaline doses




