---
title: "scratch_results"
author: "Steve Harris"
date:	"7 Mar 2016"
output: md_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls(all=TRUE))
source(file="../prep/load.R")
load(file='../data/cleaned.Rdata')
source(file="../prep/strobe.R")
source(file="../prep/prep_vars.R")
source(file="../share/functions4rmd.R")
```

# Abstract

# Introduction

Justification

- it is well known that high heart rate is associated with poor outcome 
- Morelli recently selected patients with high heart rate and randomised them to treatment with esmolol with dramatic effects,  specifically: 
	+ a reduction in noradrenaline dose
	+ a reduction in mortality
- also the question that 
	- clearly noradrenaline may be a proxy for underlying disease severity, but is it also possible that high noradrenaline is _harmful_ esp. certain subgroups i.e. those with tachycardia
- what are the deteriminants of noradrenaline dose at 24h
	- and after adjusting for these then what is the effect of noradrenaline on outcome

- 2ndry qn: apart from v high control group mortality which raised the question of practise and case mix differences 

Qns.

1. With reference to Morelli, if we capture a similar population then is the mortality observed similar to the control group - with particular relevance to patients with high heart rates
2. How much does treatment vary between sites wrt norad, fluid, and map?
3. What determines mortality at 24h, and specifically
	- what is the site effect
	- what is the effect of high heart rate
3. How much does mortality vary between sites?

# Methods

## Patients

We recruited one hundred consecutive admissions with septic shock from eight critical care units across Europe. Patients were eligible if they were locally diagnosed and treated for infection, required a vasopressor infusion after 24 hours of stabilisation in the critical care unit, and had not received beta-blocker therapy during the same time period.

<!-- Morelli criteria to compare in table

- age < 18
- beta-blockers prior to randomisation
- cardiac index < 2.2 with high LAP
- significant valvular heart disease
 -->

## Definitions

- need to describe vasopressor and sedation scores
- classify high vs low map target
	- use 75 as per Asfar
- classify high vs low heart rate groups: split at [0,95) and [95,] as per Morelli

## Research questions

- effect of map (target), heart rate, and norad dose on outcome
	+ hypothesis 1: high vs low heart rates respond differently to norad?
- inter-site variablity in treatment approaches after controlling for hour 1 status
- inter-site variability in mortality after controlling for hour 1 status and ...
	+ message: mortality is high but there is significant variability between sites  even after controlling for obvious stuff

## Statistical analaysis	



# Results
```{r strobe, include=FALSE, echo=FALSE}

los.itu.0    <- ff.np(exclude.los.itu.0)
rx.betablock <- ff.np(exclude.rx.betablock)
low.ne.24    <- ff.np(exclude.ne.24)
hosp.n       <- ff.mediqr(N, data=tdt[,.N,by=hosp])
```

We screened `r nrow(wdt[include==1])` consecutive admissions with septic shock from `r length(unique(wdt$hosp))` hospitals in 4 European countries. We excluded `r los.itu.0$n[2]` patients (`r los.itu.0$p[2]`) who were discharged within 24 hours (one survivor, one non-surivor); `r rx.betablock$n[2]` patients (`r rx.betablock$p[2]`) who received beta-blocker therapy during the first 24 hours; and, `r low.ne.24$n[2]` patients (`r low.ne.24$p[2]`) where noradrenaline infusions had been weaned by 24 hours. The final cohort included `r nrow(tdt)` patients with each hospital contributing a median of `r hosp.n$q50` patients (IQR `r hosp.n$iqr`).

## Patient characteristics

```{r patients, include=FALSE, echo=FALSE}
# source(file="../analysis/table1_all.R")
(lac24 <- ff.mediqr(lac.24,dp=1))
ff.np
(lac24gt2 <- ff.np(var=lac24gt2, data=wdt[,.(lac24gt2=ifelse(lac.24>2 & !is.na(lac.24),1,0))], dp=0)) 
```
<!-- see Table 1 -->

Vasopressor support for septic shock commenced on the first ICU day for 311 (75%) patients --- the remaining patients developing septic shock during their critical care stay. Pneumonia was the commonest cause of infection (223 patients, 46%). The median SOFA score at 24 hours was 10 (IQR 8 to 12), and lactate at 24 hours was `r lac24$q50` mmol/l (IQR `r lac24$iqr`) with `r lac24gt2$n[2]` (`r lac24gt2$p[2]`) patients have a lactate greater than 2 mmol/l. The majority of patients (424, 87%) were mechanically ventilated, and 140 (29%) had already started renal replacement therapy (RRT) at 24 hours. The overall ICU and hospital mortalities were 41% (198 deaths), and 48% (233 deaths) respectively. 


## Variation in practice between sites

- [ ] TODO(2016-03-10): remove patient physiology, report treatment only except blood pressure which you can consider a target; therefore add in sedation and vadi
- [ ] TODO(2016-03-15): add in a test for trend?

We plotted the distribution of key haemodynamic parameters using separate box and whisker plots for each critical care unit. We ordered the plots by ther mean ICU mortality for the each hospital (Figure 1). Inspection of these showed that mean arterial pressure and noradrenaline doses were higher in units that had higher mean mortalities. Conversely, first 24 hour fluid balances were higher in units with lower mortalities. 

<!-- risk: ecological fallacy -->

<!-- notes:

Now model this

- treatment variables in 1st 24 hours
	+ fluid balance
	+ sedation
	+ noradrenaline use
	+ target blood pressure

- adjust for
	- baseline severity of illness at admission
	- age
	- sex
	- admission diagnosis
	- sepsis by body system
	- control for current
		+ steroid use
		+ other vasopressors
- report variability
	+ site vs residual: how much is accounted for unit level practice

 -->

## Predictors of 24 hour treatment variables

Discuss model output for 
- ne.24
- fb.24
- map.24

## Variation in mortality between sites

- [ ] TODO(2016-03-10): compare mortality to that reported in JAMA paper
- [ ] TODO(2016-03-10): report intersite variation in mortailty
- [ ] TODO(2016-03-10): look for heart rate mortality interaction





# Discussion points

- [ ] report patients meeting criteria for septic shock as per Sepsis 3
	- blood pressure or norad (all)
	- lactate > 2
