---
title: "scratch_results"
author: "Steve Harris"
date:	"7 Mar 2016"
output: md_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls(all=TRUE))
source(file="../prep/load.R")
load(file='../data/cleaned.Rdata')
source(file="../share/functions4rmd.R")
```

# Methods

## Patients

We recruited one hundred consecutive admissions with septic shock from eight critical care units across Europe. Patients were eligible if they were locally diagnosed and treated for infection, required a vasopressor infusion (equivalent to noradrenaline at 0.1 mcg/kg/min or greater) after 24 hours of stabilisation in the critical care unit, and had not received beta-blocker therapy during the same time period.

```{r strobe, include=FALSE, echo=FALSE}
wdt[, include := ifelse(age>=18 & ne.1 > 0, 1,0) ]
wdt[, include := ifelse(is.na(include),0,include)]
describe(wdt$include)
wdt <- wdt[include==1]

wdt[, exclude.los.itu.0 := ifelse(los.itu<=0,1,0)]
with(wdt[los.itu<=7], table(los.itu, mort.itu))
los.itu.0 <- ff.np(exclude.los.itu.0)
describe(wdt$ne.24)
wdt[, exclude.rx.betablock := ifelse(rx.betablock==FALSE | is.na(rx.betablock),0,1)]
rx.betablock <- ff.np(exclude.rx.betablock)
wdt[, exclude.ne.24 := ifelse(ne.24 >= 0.1 & !is.na(ne.24),0,1)]
low.ne.24 <- ff.np(exclude.ne.24)
```

# Results

We screened `r nrow(wdt[include==1])` consecutive admissions with septic shock from `r length(unique(wdt$hosp))` hospitals in 4 European countries. We excluded `r los.itu.0$n[2]` patients (`r los.itu.0$p[2]`) who were discharged within 24 hours (one survivor, one non-surivor); `r rx.betablock$n[2]` patients (`r rx.betablock$p[2]`) who received beta-blocker therapy during the first 24 hours; and, `r low.ne.24$n[2]` patients (`r low.ne.24$p[2]`) where the noradrenaline dose had been weaned to below 0.1 mcg/kg/min at 24 hours.

## Patient characteristics

```{r patients, include=FALSE, echo=FALSE}
tdt <- wdt[include==1 & exclude.los.itu.0==0 & exclude.rx.betablock==0 & exclude.ne.24 ==0]
nrow(tdt)


```


see Table 1


Morelli criteria to compare in table

- age < 18
- beta-blockers prior to randomisation
- cardiac index < 2.2 with high LAP
- significant valvular heart disease

## Variation in practice between sites

- sedation
- fluid management
- patient characteristics

## Variation in mortality between sites

a
## Model 24 hour noradrenaline doses




