---
title: "scratch_results"
author: "Steve Harris"
date:	"3 May 2016"
output: word_document
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls(all=TRUE))
source(file="../prep/strobe.R")
source(file="../prep/prep_vars.R")
source(file="../share/functions4rmd.R")
source(file="../share/functions4paper.R")
```

```{r strobe, include=FALSE, echo=FALSE}

(los.itu.0    <- ff.np(data=wdt.original, exclude.los.itu.0))
with(wdt.original, table(exclude.los.itu.0, is.na(ne.1)))
with(wdt.original, table(exclude.los.itu.0, mort.itu))
(include    <- ff.np(data=wdt.original, include))
(grp.ne24    <- ff.np(data=wdt, grp.ne24))
(rx.betablock <- ff.np(data=wdt.original, exclude.rx.betablock))
(grp.morelli    <- ff.np(data=wdt, grp.morelli))
(hosp.n       <- ff.mediqr(N, data=wdt[,.N,by=hosp]))
```

We screened `r nrow(wdt.original)` consecutive admissions with septic shock from `r length(unique(wdt$hosp))` hospitals in 4 European countries. We excluded 45 patients (6 missing data on their initial noradrenaline requirement, and 39 deaths within 24 hours). Each hospital contributing a median of `r hosp.n$q50` patients (IQR `r hosp.n$iqr`).

The subgroup of patients with resistant septic shock comprised `r grp.ne24$n[2]` patients (`r grp.ne24$n[2]`) who had not been weaned from noradrenaline within 24 hours. Within this, the tachycardic-shocked subgroup included `r grp.morelli$n[2]` (`r grp.morelli$p[2]`) patients with a heart rate greater than 95 beats/minute without preceding beta-blocker treatment. Subgroups were not evenly distributed between ICUs (Figure 1).

## Patient characteristics

```{r patients, include=FALSE, echo=FALSE}
# Need to inspect Table 1

(sofa1 <- ff.mediqr(sofa.1,dp=1))
(lac24 <- ff.mediqr(lac.24,dp=1))
(lac24gt2 <- ff.np(var=lac24gt2, data=wdt[,.(lac24gt2=ifelse(lac.24>2 & !is.na(lac.24),1,0))], dp=0)) 
```
<!-- see Table 1 -->

Septic shock occurred within 48 hours of admission for 460 (80%) of patient. Pneumonia was the commonest cause (292 patients, 42%). Patient variables at 24 hours following onset of septic shock are reported in Table 1. The median SOFA score was 10 (IQR 7 to 12), and the lactate at the same time was `r lac24$q50` mmol/l (IQR `r lac24$iqr`) with `r lac24gt2$n[2]` (`r lac24gt2$p[2]`) patients have a lactate greater than 2 mmol/l. The majority of patients (556, 77%) were mechanically ventilated, and 178 (25%) had already started renal replacement therapy (RRT). The overall ICU mortality was 35% (242 deaths). For the subgroups with resistant shock, and tachycardic-shock the ICU mortalities were 40% (227 deaths), and 49% (131 deaths respectively).

## Treatment variation between ICUs
```{r treatment variation, include=FALSE, echo=FALSE}
source("../analysis/fig_process_var.R")

t1.trend("ne.24", "hosp.id.sort", data=wdt, var.cont=TRUE)
t1.trend("map.24", "hosp.id.sort", data=wdt, var.cont=TRUE)
t1.trend("fb.24", "hosp.id.sort", data=wdt, var.cont=TRUE)

wdt[, rx.roids.01 := ifelse(rx.roids==TRUE, 1, 0)]
wdt[, .(mean(rx.roids.01, na.rm=TRUE), mean(mort.itu), .N), by=hosp.id.sort]
t1.trend("rx.roids.01", "hosp.id.sort", data=wdt, var.cont=FALSE)
t1.trend("rx.roids.01", "hosp.id.sort", data=wdt, var.cont=FALSE)

t1.trend("rrt.24", "hosp.id.sort", data=wdt, var.cont=FALSE)
t1.trend("mv.24", "hosp.id.sort", data=wdt, var.cont=FALSE)
# - [ ] TODO(2016-03-25): add in test for trend for VADI/Sedation/RRT

```

We examined noradrenaline doses administered at 24 hours after ranking the ICUs by their mean 24 hour mortality (Figure 2, top panel). Noradrenaline doses were higher in units that had higher mean mortalites (test for trend by comparing the coefficient in a linear model to zero, p<0.0001). We repeated the process for the treatment variables that might affect 24 hour noradrenaline dose (mean arterial blood pressure, fluid balance, other vasoactive drugs, steroids, sedation, and other organ support). This showed that mean arterial pressure (p=0.015) was higher (p=0.014), and fluid balances lower in units with higher mean mortality. Similarly, other vasoactive drugs (p<0.001), corticosteroids (p=0.01) and sedation (p<0.001) were used more often, as was mechanical ventilation (p=0.0001), but not renal replacement therapy (p=0.056) (see supplementary Figure 1).

## Predictors of 24 hour treatment variables

For noradrenaline, mean arterial pressure and fluid balance, we entered each variable into a model to estimate the relative contributions of patient predictors (at baseline) and the ICU level effect on these measures at 24 hours (Figure 2 [lower panel], and Table 2). For example, higher baseline SOFA scores and lactate measures are were associated with noradrenaline requirement at 24 hours. However, the same variables had little effect on MAP at the same time point (presumably because it was defined by noradrenaline dose instead), and baseline lactate but not the SOFA score was correlated with fluid balance.

The ICC (proportion of variance in the multilevel model accounted for by the ICU) for the 24 hour values of noradrenaline dose was 44%, for MAP was 4%, and for fluid balance was 29%.

## Mortality and noradrenaline dose at 24 hours

The final ICU mortality model demonstrated the importance of noradrenaline dose at 24 hours as a predictor of survival (OR XX, Figure 3 and Table ). Mean arterial pressure after adjustment for other variables had no effect on outcome (OR XX). Increasing heart rate at 24 hours was associated with higher mortality (OR XX), but this effect was attenuated as heart rate increased (Figure 4). The ICC for this model was 7% indicating that the majority of the variation in outcoome arose from differences in patient characteristics.


<!-- - [ ] TODO(2016-04-30): convert linear estimates to exponentials for OR
- [ ] TODO(2016-03-10): compare mortality to that reported in JAMA paper
- [ ] TODO(2016-03-10): report intersite variation in mortailty
- [ ] TODO(2016-03-10): look for heart rate mortality interaction
 -->







